Body size patterns

### 2019-03-28 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
added spp-specific cohortStrength effects

cohortStrength <- vB %>%
  distinct(tagIndex, species, metaPop01, cohort) %>% #ignoring 10 fish that were in both o'bear and WB
  group_by(species, metaPop01, cohort) %>%
  summarize(cohortStrength = n()) %>%
  filter(!(cohort <2000 & species !='ats')) %>% # no trib data
  mutate(cohortStrengthScaled = scale(cohortStrength))

@@

@@

AIC(mod,modC,modR,modC_R,modCR, modCR_c, modCR_S_c, modCR_c_cS, modCR_S_c_cS, modCR_SR_c_cS, modCR_SR_c_cSR) %>% rownames_to_column() %>% arrange(AIC)

@@

@@

# Brook trout

modBKT <- bam(observedLength ~ cohort * river + factor(salmon01) +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")

@@

modBNT <- bam(observedLength ~ cohort * river + factor(salmon01) +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")

@@

# specific splines by species
plotCohort <- plotGamVarBySpecies(var = 's(cohort)', xAxisLab = 'Cohort')

>>>>

cohortStrength <- vB %>%
  distinct(tagIndex, species, metaPop01, cohort) %>% #ignoring 10 fish that were in both o'bear and WB
  group_by(species, metaPop01, cohort) %>%
  summarize(cohortStrength = n()) %>%
  filter(!(cohort < 2000 & species !='ats')) %>% # no trib data
  mutate(cohortStrengthScaled = scale(cohortStrength))

@@

cohortStrengthWide <- spread(cohortStrength, species, cohortStrengthScaled) %>%
  rename(cohortStrengthScaledBKT = bkt,
         cohortStrengthScaledBNT = bnt,
         cohortStrengthScaledATS = ats
         ) %>%
  dplyr::select(-cohortStrength)

vB <- left_join(vB, cohortStrengthWide)

vB$cohortStrengthScaledBKT <- ifelse(is.na(vB$cohortStrengthScaledBKT), -2.5, vB$cohortStrengthScaledBKT)
vB$cohortStrengthScaledBNT <- ifelse(is.na(vB$cohortStrengthScaledBNT), -2.5, vB$cohortStrengthScaledBNT)
vB$cohortStrengthScaledATS <- ifelse(is.na(vB$cohortStrengthScaledATS), -2.5, vB$cohortStrengthScaledATS)

@@

# effects of each species together - above is only conspecific
system.time(
modCR_SR_c_cSR3 <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverOrdered) +
               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
             data = dataIn, method = "REML")
)
summary(modCR_SR_c_cSR3)
#plot(modCR_SR_c_cSR3, scale = 0, pages=5)

system.time(
modCR_SR_c_cSR3a <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn) +
               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
             data = dataIn, method = "REML")
)
summary(modCR_SR_c_cSR3a)
#plot(modCR_SR_c_cSR3a, scale = 0, pages=5)

system.time(
modCR_SR_c_cSR3b <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = metaPop01) +
               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
             data = dataIn, method = "REML")
)
summary(modCR_SR_c_cSR3b)
#plot(modCR_SR_c_cSR3b, scale = 0, pages=5)

AIC(mod,modC,modR,modC_R,modCR, modCR_c, modCR_S_c, modCR_c_cS, modCR_S_c_cS, modCR_SR_c_cS, modCR_SR_c_cSR, modCR_SR_c_cSR3, modCR_SR_c_cSR3a,modCR_SR_c_cSR3b) %>% rownames_to_column() %>% arrange(AIC)

@@

# BKT

# BNT
#             rowname       df      AIC
# 1  modCR_SR_c_cSR3a 89.52290 231873.4
# 2  modCR_SR_c_cSR3b 89.54674 231873.6
# 3     modCR_SR_c_cS 84.40505 231935.7
# 4      modCR_S_c_cS 83.47382 231962.5
# 5   modCR_SR_c_cSR3 88.15840 232100.8
# 6        modCR_c_cS 80.92171 232142.0
# 7    modCR_SR_c_cSR 80.89406 232201.5
# 8         modCR_S_c 79.53048 232281.7
# 9           modCR_c 77.80808 232420.2
# 10            modCR 78.58400 233816.0
# 11           modC_R 53.21199 234214.3
# 12              mod 13.99759 238544.8
# 13             modR 14.97078 243472.0
# 14             modC 48.73277 431440.2

@@

# Brook trout
# using salmonid01 * river because we need diff effects per river - even though AIC is worse for this model 
modBKT <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")

@@

modBNT <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = metaPop01) +
               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")

@@

# specific splines by species
(plotCohort <- plotGamVarBySpecies(var = 's(cohort)', xAxisLab = 'Cohort'))
(plotCohortStrength <- plotGamVarBySpecies(var = 's(cohortStrengthScaled)', xAxisLab = 'Cohort Strength'))

#
